appendix

### 2019-09-12 manuscript/appendix.Rmd
Create appendix.Rmd

---
title: "Manuscript draft"
author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    includes:
      after_body: appendix.md
bibliography: [library.bib,  packages.bib]
biblio-style: apalike
csl: apa.csl
header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}

---

@@

```{r include=FALSE}
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
), 'packages.bib')
```

```{r results="asis", include=FALSE}
cat("
<style>
caption {
      font-size: 0.75em;
    }
</style>
")
```

###Abstract

Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle. Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes. Furthermore, the climate variables that drive this relationship are poorly understood. Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines. The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit. Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components. Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity. However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components. Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.

### Introduction

Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate. Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**). Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005 (my first paper!!), but check more recent literature**). On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**). The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**). 
Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme. However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time. The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.

On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011]. Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011]. However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011]. This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007]. In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017]. Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
An understanding of the global patterns of productivity is important; however elucidating the drivers of these patterns will be even more valuable in expanding our understanding of global carbon cycling. Primary productivity can be influenced by many factors, which often act across a range of scales, and may show interactive effects with each other. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity. On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.

We know from previous research that climate is a significant driver of productivity across broad spatial scales [@cleveland_relationships_2011]. The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate. There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016]. However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies. There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003]. Studies of the influence of MAT on productivity are less conclusive. Luyssaert et al. [-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT. In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature. The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them. Taylor et al. [-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.

While MAT and MAP are important climate variables, they do not capture all aspects of climate that may influence productivity, and may be insufficient to capture the breadth of climatic effects on productivity [@cleveland_relationships_2011]. There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways. Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a factor of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, indicating that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.

In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1). Firstly, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables. We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season.

Resolving these questions requires data which, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes. Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.

\renewcommand{\arraystretch}{2}
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
hypothesis_table <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(hypothesis_table, booktabs = TRUE, caption = "Hypothesis table") %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
  pack_rows("1. How do FACF vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("2. How do FACF relate to MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("3. How do FACF relate to other annual climate variables?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("4. What is the role of seasonality?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("5. Per month of the growing seaon, how do FACF vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
  column_spec(1, width = "10cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3:13, width = "1.8cm") %>%
  kableExtra::landscape()
```

### Materials and Methods

Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]. This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions. ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.

```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
  knitr::include_graphics("distribution_all_variables.png")
```
*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity. Table 1 contains details of the variables selected for analysis.

| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
|--------------|-------------------|-------------|------------------------|-------|-------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 92 | 42 |
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |

Table: Definitions and sample sizes of variables used in analysis. Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.

A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors. Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. Sites that had experienced significant disturbance were also excluded. Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees. Grazed sites were retained.  
There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015]. To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis. Sites for which stand age was unknown were excluded from analysis.

*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods. For this reason, criteria were introduced to standardise for differences in methodology. Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.  It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.

As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts. Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis. Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root. Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded. Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.

*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database. Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v. 4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables). Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.

Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.

*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. Growing season months were initially calculated following methods used by Kerkhoff et al. (2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95. Michaletz et al. (2014) included an equivalent requirement in their calculation of growing season length. However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.

Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season. Total growing season precipitation and solar radiation were also calculated.

*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect. For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term. AIC values were calculated for the models and used to select the best model. If the best model included a polynomial term, the shape of the polynomial relationship was considered. If the shape of the relationship made biological sense, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model. If not, we ran the linear model as the final model. R^2^ values were calculated for the best model. All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified. In addition, slope coefficients were calculated for the linear models.

Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1. As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.

To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect. These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models. Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power. As a result, it was decided to present results only from models do not include altitude as a fixed term.

Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas. To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area. Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest. These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone. Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.

Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models. Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature. We therefore decided to include mean annual temperature as a term in all multivariate models. We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability. Models were tested for a significant interactive effect and a significant additive effect. We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation. In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate. For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. As before, plot nested within geographic area was included as a random effect. Altitude was not considered. AIC values were calculated for the models, and used to compare models. Models were considered to be significantly better than the baseline MAT*MAP model if:  
  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5

*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes. To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component. The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. Stacked plots were generated for:  
1. GPP = NPP + R~auto~  
2. NPP = ANPP + BNPP  
3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
4. Total belowground carbon flux = BNPP + R~root~

*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~. For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated. The ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions. To account for this, models were re-run using only data from sites $\leq$ 1000m.

###Results

In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 154 distinct geographic areas (Fig. 1, Table 2), across all forested biogeographic and climate zones.

```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$. Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes. 95\\% confidence intervals are plotted for the larger flux. Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}

  knitr::include_graphics("combined_stacked.png")
```

*How does productivity vary with latitude?*

All major carbon fluxes increased linearly with decreasing latitude (fig. 3).

```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("effect_of_lat_transparent.png")
```

Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). For all fluxes, their relationship with latitude was best predicted by the linear model.

*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig. 2). That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

We found no evidence that allocation between fluxes varied substantially with latitude or climate. There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.

R^2^ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2). Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~. The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP. Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower. Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~. In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation. This pattern was also seen in the R^2^ values for multivariate models.

*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP. MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables. MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig. 5). We found no support for a saturation point of productivity with temperature.

MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux. For the majority of fluxes productivity was best predicted by a polynomial model. Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig. 5). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33. Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone. However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation. Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was evidence that productivity begins to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for all fluxes. BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 5). Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, productivity began to decrease.

All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation. Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).

Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("combined_plots.png")
```

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig. 5). ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it. Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

In contrast, there was no significant effect of precipitation seasonality on productivity.

We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig. 5). Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.

*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity. For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes. There was a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05). Productivity increased with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).

###Discussion

In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale. Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data. This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.

Our results show that productivity decreases linearly with latitude. Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale. While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically. Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP. There is a significant interaction between MAT and MAP.  We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.

We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor. Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.

Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses. We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux. However, field techniques for estimating components of primary productivity are variable in their accuracy. This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP. Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux. It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality. While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP. We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change. It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.

Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale. Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites. As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases. While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity. Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT. With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.

Furthermore, we expect that temporal resolution of climate data affects our results. Productivity is known to vary on a seasonal and annual scale in response to variation in climate. Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation. Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked. The limitations of low temporal resolution are highlighted in our analysis of growing season length. We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages. With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.

Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.

>>>>

---
title: "appendix"
output: 
  pdf_document:
    fig_caption: yes

header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
---

@@
